---
title: "Benchmark : matrix; data.frame; data.table"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

matrix data.frame data.table

  * Row appending
    * Good performance : data.table(rbindlist) ; data.table(rbind) ; matrix(rbind)
  * Row deletion
    * Good performance : matrix([,-1]) ; matrix(tail) ; data.table([,-1]) ; data.table(tail) ; 
  * Row selection
    * Good performance : data.frame([,]) ; matrix([,]) ; data.table(subset)
    
 

Summary : Matrix and data.table are very close in performance. data.frame should be avoid in data and operation intensive cases 

# Benchmark

## Row appending  

```{r, warning=FALSE}
library(microbenchmark) # for microbenchmark
library(data.table)

#----------

mat0 <- matrix(ncol = 100, nrow = 1000)
df0 <- data.frame(mat0)
dt0 <- data.table(mat0)

newRow <- 1:100
loops <- 1:5000 # ranking result may different depending on the loop size

# ---------

# rbind
mat_rbind <- function(){
  mat <- mat0
  for (i in loops) {
    mat <- rbind(mat, newRow)
  }
}

# rbind
df_rbind <- function(){
  df <- df0
  for (i in loops) {
    df <- rbind(df, newRow)
  }
}

# insert
df_insert <- function(){
  df <- df0
  for (i in loops) {
    df[nrow(df)+1,] <- newRow
  }
}

# rbind
dt_rbind <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- rbind(dt, as.list(newRow))
  }
}

# rbindlist
dt_rbindlist <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- rbindlist(list(dt, as.list(newRow) ) )
  }
}

```

```{r, eval=FALSE}
res <- microbenchmark(
  mat_rbind()
  , df_rbind()
  , df_insert()
  , dt_rbind()
  , dt_rbindlist()
  , times = 50
)

print(res)

```

```r
Unit: seconds
     expr      min       lq     mean   median       uq       max neval  cld
   mat_rbind() 29.61510 29.91575 30.20403 30.08436 30.20831  35.67312    50  b  
    df_rbind() 87.38882 88.11399 89.57409 88.34435 88.73782 112.14921    50    d
   df_insert() 48.18800 48.56083 49.34227 48.84676 49.45061  58.04525    50   c 
    dt_rbind() 25.22989 25.42573 25.73572 25.53962 25.71169  33.92170    50 a   
dt_rbindlist() 24.56441 24.74358 25.47515 24.81044 25.02734  37.53949    50 a  
```

## Row deletion

```{r}
mat0 <- matrix(ncol = 100, nrow = 10000)
df0 <- data.frame(mat0)
dt0 <- data.table(mat0)

loops <- 1:5000

#-----------------------

mat_delete <- function(){
  mat <- mat0
  for (i in loops) {
    mat <- mat[-1, ]
  }
}

mat_tail <- function(){
  mat <- mat0
  for (i in loops) {
    mat <- tail(mat, nrow(mat)-1 )
  }
}

df_delete <- function(){
  df <- df0
  for (i in loops) {
    df <- df[-1, ]
  }
}

df_tail <- function(){
  df <- df0
  for (i in loops) {
    df <- tail(df, nrow(df)-1 )
  }
}

dt_delete <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- dt[-1, ]
  }
}

dt_tail <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- tail(dt, nrow(dt)-1 )
  }
}
```


```{r, eval=FALSE}
res <- microbenchmark(
  mat_delete()
  , mat_tail()
  , df_delete()
  , df_tail()
  , dt_delete()
  , dt_tail()
  , times = 50
)

print(res)
```

```r
## Unit: seconds
##          expr      min       lq     mean   median       uq      max neval 
##  mat_delete() 34.70892 35.27614 35.58425 35.43062 35.67061 38.85541    50  a
##    mat_tail() 38.09419 38.92384 39.26626 39.12545 39.33041 42.16310    50   b
##   df_delete() 47.66720 47.85000 48.04832 47.99169 48.12198 49.68098    50
##     df_tail() 74.40627 75.18312 75.42762 75.40833 75.54480 76.55802    50
##   dt_delete() 40.48840 40.72822 40.85509 40.87139 40.90501 41.73669    50
##     dt_tail() 41.30999 41.58012 41.68345 41.64761 41.75733 42.54702    50

```

##  Row selection

```{r}
mat0 <- matrix(ncol = 50, nrow = 10000)
mat0[, 1] <- 1:10000 
mat0[, 2] <- 1:10000 

df0 <- data.frame(mat0)
dt0 <- data.table(mat0)

loops <- 1000:6000 # loop 5k times
#----------------------

mat_select <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(mat0[mat0[,1]>i & mat0[,1]<i+50, 2]) + sum
  }
  print(sum)
}

mat_subset <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(subset(mat0[,2], mat0[,1]>i & mat0[,1]<i+50)) + sum
  }
  print(sum)
}

df_select <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(df0[df0[,1]>i & df0[,1]<i+50, 2]) + sum
  }
  print(sum)
}

df_subset <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(subset(df0[,2], df0[,1]>i & df0[,1]<i+50)) + sum
  }
  print(sum)
}

dt_select <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(dt0[dt0$V1>i & dt0$V1<i+50, "V2", with=FALSE])+sum
  }
  print(sum)
}

dt_subset <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(subset(dt0$V2, dt0$V1>i & dt0$V1<i+50)) + sum
  }
  print(sum)
}
```

```{r, eval=F}
res <- microbenchmark(
  mat_select()
  , mat_subset()
  , df_select()
  , df_subset()
  , dt_select()
  , dt_subset()
  , times = 50
)

print(res)
```

```r
## Unit: seconds
##          expr      min       lq     mean   median       uq      max neval
##  mat_select() 2.651344 2.684528 2.705679 2.697216 2.709088 2.808091    50
##  mat_subset() 4.506899 4.537437 4.591004 4.563015 4.608267 4.892726    50
##   df_select() 1.943527 1.994755 2.015304 2.011318 2.040349 2.116544    50
##   df_subset() 3.183609 3.245042 3.290002 3.263126 3.299738 3.546379    50
##   dt_select() 3.937172 4.019212 4.072775 4.048270 4.090016 4.553862    50
##   dt_subset() 3.164424 3.220253 3.278391 3.259319 3.311617 3.694738    50
##    cld
##   b   
##      e
##  a    
##    c  
##     d 
##    c

```




