---
title: "Benchmark : Matrix data.frame data.table"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

Matrix data.frame data.table

  * Row appending
    * good performance : data.table(rbindlist) ; data.table(rbind) ; matrix(rbind)
  * Row deletion
    * good performance : matrix([,-1]) ; matrix(tail) ; data.table([,-1]) ; data.table(tail) ; 
  * Row selection


Summary : Matrix and data.table are very close in performance. data.frame should be avoid in data and operation intensive cases 


# Row appending  

```{r, warning=FALSE}
library(microbenchmark) # for microbenchmark
library(data.table)

#----------

mat0 <- matrix(ncol = 100, nrow = 1000)
df0 <- data.frame(mat0)
dt0 <- data.table(mat0)

newRow <- 1:100
loops <- 1:5000 # ranking result may different depending on the loop size

# ---------

# rbind
mat_rbind <- function(){
  mat <- mat0
  for (i in loops) {
    mat <- rbind(mat, newRow)
  }
}

# rbind
df_rbind <- function(){
  df <- df0
  for (i in loops) {
    df <- rbind(df, newRow)
  }
}

# insert
df_insert <- function(){
  df <- df0
  for (i in loops) {
    df[nrow(df)+1,] <- newRow
  }
}

# rbind
dt_rbind <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- rbind(dt, as.list(newRow))
  }
}

# rbindlist
dt_rbindlist <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- rbindlist(list(dt, as.list(newRow) ) )
  }
}

```

```{r, eval=FALSE}
res <- microbenchmark(
  mat_rbind()
  , df_rbind()
  , df_insert()
  , dt_rbind()
  , dt_rbindlist()
  , times = 50
)

print(res)

```

```r
Unit: seconds
     expr      min       lq     mean   median       uq       max neval  cld
   mat_rbind() 29.61510 29.91575 30.20403 30.08436 30.20831  35.67312    50  b  
    df_rbind() 87.38882 88.11399 89.57409 88.34435 88.73782 112.14921    50    d
   df_insert() 48.18800 48.56083 49.34227 48.84676 49.45061  58.04525    50   c 
    dt_rbind() 25.22989 25.42573 25.73572 25.53962 25.71169  33.92170    50 a   
dt_rbindlist() 24.56441 24.74358 25.47515 24.81044 25.02734  37.53949    50 a  
```

# Row deletion

```{r}
mat0 <- matrix(ncol = 100, nrow = 10000)
df0 <- data.frame(mat0)
dt0 <- data.table(mat0)

loops <- 1:5000

#-----------------------

mat_delete <- function(){
  mat <- mat0
  for (i in loops) {
    mat <- mat[-1, ]
  }
}

mat_tail <- function(){
  mat <- mat0
  for (i in loops) {
    mat <- tail(mat, nrow(mat)-1 )
  }
}

df_delete <- function(){
  df <- df0
  for (i in loops) {
    df <- df[-1, ]
  }
}

df_tail <- function(){
  df <- df0
  for (i in loops) {
    df <- tail(df, nrow(df)-1 )
  }
}

dt_delete <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- dt[-1, ]
  }
}

dt_tail <- function(){
  dt <- dt0
  for (i in loops) {
    dt <- tail(dt, nrow(dt)-1 )
  }
}
```


```{r, eval=FALSE}
res <- microbenchmark(
  mat_delete()
  , mat_tail()
  , df_delete()
  , df_tail()
  , dt_delete()
  , dt_tail()
  , times = 50
)

print(res)
```

```r
## Unit: seconds
##          expr      min       lq     mean   median       uq      max neval 
##  mat_delete() 34.70892 35.27614 35.58425 35.43062 35.67061 38.85541    50  a
##    mat_tail() 38.09419 38.92384 39.26626 39.12545 39.33041 42.16310    50   b
##   df_delete() 47.66720 47.85000 48.04832 47.99169 48.12198 49.68098    50
##     df_tail() 74.40627 75.18312 75.42762 75.40833 75.54480 76.55802    50
##   dt_delete() 40.48840 40.72822 40.85509 40.87139 40.90501 41.73669    50
##     dt_tail() 41.30999 41.58012 41.68345 41.64761 41.75733 42.54702    50

```

#  Row selection

```{r}
mat0 <- matrix(ncol = 50, nrow = 10000)
mat0[, 1] <- 1:10000 
mat0[, 2] <- 1:10000 

df0 <- data.frame(mat0)
dt0 <- data.table(mat0)

loops <- 1000:6000

#----------------------

mat_select <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(mat0[mat0[,1]>i & mat0[,1]<i+50, 2]) + sum
  }
  print(sum)
}

mat_subset <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(subset(mat0[,2], mat0[,1]>i & mat0[,1]<i+50)) + sum
  }
  print(sum)
}

df_select <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(df0[df0[,1]>i & df0[,1]<i+50, 2]) + sum
  }
  print(sum)
}

df_subset <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(subset(df0[,2], df0[,1]>i & df0[,1]<i+50)) + sum
  }
  print(sum)
}

dt_select <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(dt0[dt0$V1>i & dt0$V1<i+50, "V2", with=FALSE])+sum
  }
  print(sum)
}

dt_subset <- function(){
  sum <- 0
  for (i in loops) {
    sum <- sum(subset(dt0$V2, dt0$V1>i & dt0$V1<i+50)) + sum
  }
  print(sum)
}
```

```{r, eval=FALSE}
res <- microbenchmark(
  mat_select()
  , mat_subset()
  , df_select()
  , df_subset()
  , dt_select()
  , dt_subset()
  , times = 50
)

print(res)
```

```r
## Unit: nanoseconds
##          expr        min         lq         mean     median         uq
##    mat_select          0          1 3.768200e+02        428        429
##  mat_subset() 4276422959 4336275217 4.805567e+09 5021651896 5114309661
##   df_select() 1872613721 1922426744 2.155157e+09 2233557692 2326910339
##   df_subset() 3046043127 3129437574 3.525092e+09 3650319753 3734093072
##   dt_select() 3703176576 3764355310 4.491682e+09 4733422992 4825856470
##   dt_subset() 3015270310 3119553566 3.526231e+09 3629489957 3740920881
##         max neval   cld
##        1711    50 a    
##  5530042228    50     e
##  2659248458    50  b   
##  4043180638    50   c  
##  5091721065    50    d 
##  3949492095    50   c
```




