---
title: "Reduction into a single variable/dimension"
date: "Updated: `r format(Sys.time(), '%d %B, %Y')`"
---

# Data compression

Reduce mulitple variables/dimensions into a single variable/dimension.

Three different approaches: Mean, Median, and First principle component

## build the dataset

```{r}
library(scales)

N <- 10 # number of variables
x1 <- sample(seq(1, 20, by=1), N, replace = T)
x2 <- sample(seq(1, 150, by=1), N, replace = T)
x3 <- rnorm(N, mean=50, sd=10) 
x4 <- rnorm(N, mean=50, sd=30)
x5 <- dlnorm(1:N, 2, 0.35)    # log distribution
x5 <- rescale(x5  , to=c(0,150))
df <- data.frame(x1,x2,x3,x4,x5)
df <- t(df)
df <- round(df, digits=2)
```

dataset contains 5 instances, each instance contains 10 variables

```{r}
df
```


## plot data with comparsion

plot all data points

```{r}
# plot  
plot(c(1,1), xlim=c(1, nrow(df)), ylim=c(min(df), max(df)), type = 'n')
for(i in 1:nrow(df)){
  points(array(i, ncol(df)), df[i,]) # plot each row
}
```

plot data with comparsion

```{r}
# box and whisker plot
boxplot(t(df), main="blue=mean, black_bar=median, red=1st pc")

# plot average 
row_mean <- apply(df, 1, mean)
lines(x=c(1: nrow(df)), y=row_mean, col="blue")

# plot first principle component in original scale
pca <- prcomp(df) 
# reverse PCA in prcomp to get original data
# prcomp will center the variables so you need to add the subtracted means back
ori <- t(t(pca$x %*% t(pca$rotation)) + pca$center)
# first column of ori to get origional values based on first principle component
lines(x=c(1: nrow(df)), y=ori[, 1], col="red") 
```

For Reducing mutiple variables into a single variable, principle component (1st) is not good comparing to mean and median.
While, PCA can reduce mutiple variables into more than one variables
