---
title: "Densitty estimation"
output: html_document
date: "Updated: `r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Denisty estimation

http://www.wikiwand.com/en/Density_estimation#/Example_of_density_estimation

```{r}

library(MASS)
data(Pima.tr)
data(Pima.te)

Pima <- rbind (Pima.tr, Pima.te)
glu  <- Pima[, 'glu']

# data example
Pima [1:7, c('glu','type')]


d0 <- Pima[, 'type'] == 'No'
d1 <- Pima[, 'type'] == 'Yes'

glu.density    <- density (glu)
glu.d0.density <- density (glu[d0])
glu.d1.density <- density (glu[d1])


# plot denstiy function
plot(density(glu[d0]), col='blue', xlab='glu'
       , main='estimate p(glu|not diabetes) blue, p(glu|diabetes) red, p(glu) black')
lines(density(glu[d1]), col='red')
lines(density(glu), col='black')

```


# estimate probability of diabete conditional on glu 

From the density of "glu" conditional on diabetes, we can obtain the probability of diabetes conditional on "glu" via Bayes' rule. 

p.d.given.glu function is for

$$
p(diabetes=1|glu)=
\frac{p(glu|diabetes=1)*p(diabetes=1)}
{p(glu|diabetes=1)p(diabetes=1)+p(glu|diabetes=0)p(diabetes=0)}
$$

```{r}
base.rate.d1 <- sum(d1) / (sum(d1) + sum(d0))

# approxfun - linear (or constant) interpolation Functions
approxfun(glu.d0.density$x, glu.d0.density$y) -> glu.d0.f
approxfun(glu.d1.density$x, glu.d1.density$y) -> glu.d1.f

p.d.given.glu <- function(glu, base.rate.d1)
{
  p1 <- glu.d1.f(glu) * base.rate.d1
  p0 <- glu.d0.f(glu) * (1 - base.rate.d1)
  p1 / (p0 + p1)
}

x <- 1:250
y <- p.d.given.glu (x, base.rate.d1)
plot(x, y, type='l', col='red', xlab='glu', ylab='estimated p(diabetes|glu)')

```

figure shows the estimated posterior probability p(diabetes=1 | glu). From these data, it appears that an increased level of "glu" is associated with diabetes.


