---
title: "MCMC"

---


# MCMC

y = a*x + b + sd (standard deviation of the error)

create a sample dataset y = 5x + 0 + 10(sd)
```{r}
trueA <- 5
trueB <- 0
trueSd <- 10
sampleSize <- 31
 
# create independent x-values 
x <- (-(sampleSize-1)/2):((sampleSize-1)/2)
# create dependent values according to ax + b + N(0,sd)
y <-  trueA * x + trueB + rnorm(n=sampleSize, mean=0, sd=trueSd)
 
plot(x,y, main="Test Data")
```


```{r}
likelihood <- function(param){
    a = param[1]
    b = param[2]
    sd = param[3]
    
    # simply calculate the difference between predictions y = b + a*x and the observed y, 
    # and then we  look up the probability densities (using dnorm) for such deviations to occur.
    pred = a*x + b
    singlelikelihoods = dnorm(y, mean = pred, sd = sd, log = T)
    sumll = sum(singlelikelihoods)
    return(sumll)   
}
 
# Example: plot the likelihood profile of the slope a
slopevalues <- function(x){
    return(likelihood(c(x, trueB, trueSd)))
  }
slopelikelihoods <- lapply(seq(3, 7, by=.05), slopevalues )
plot (seq(3, 7, by=.05), slopelikelihoods , type="l", xlab = "values of slope parameter a", ylab = "Log likelihood")
```

```{r}
# Prior distribution
prior <- function(param){
    a = param[1]
    b = param[2]
    sd = param[3]
    aprior = dunif(a, min=0, max=10, log = T)
    bprior = dnorm(b, sd = 5, log = T)
    sdprior = dunif(sd, min=0, max=30, log = T)
    return(aprior+bprior+sdprior)
}
```

posterior = prior * likeihood

the sum because we work with logarithms

```{r}
posterior <- function(param){
   return (likelihood(param) + prior(param))
}
```

Metropolis algorithm

* Starting at a random parameter value
* Choosing a new parameter value close to the old value based on some probability density that is called the proposal function
* Jumping to this new point with a probability p(new)/p(old), where p is the target function, and p>1 means jumping as well


```{r}
######## Metropolis algorithm ################
 
proposalfunction <- function(param){
    return(rnorm(3, mean = param, sd= c(0.1,0.5,0.3)))
}
 
run_metropolis_MCMC <- function(startvalue, iterations){
    chain = array(dim = c(iterations+1,3))
    chain[1,] = startvalue
    for (i in 1:iterations){
        proposal = proposalfunction(chain[i,])
         
        probab = exp(posterior(proposal) - posterior(chain[i,]))
        if (runif(1) < probab){
            chain[i+1,] = proposal
        }else{
            chain[i+1,] = chain[i,]
        }
    }
    return(chain)
}
 
# randomly intianl parameter values (a, b, sd) 
# the proposalFunction will proposal a new value based on input paramer value
startvalue = c(4,0,10) 
chain = run_metropolis_MCMC(startvalue, 10000)
 
burnIn = 5000
acceptance = 1-mean(duplicated(chain[-(1:burnIn),]))
```


```{r}
### Summary: #######################
 
par(mfrow = c(2,3))
hist(chain[-(1:burnIn),1],nclass=30, main="Posterior of a", xlab="True value = red line" )
abline(v = mean(chain[-(1:burnIn),1]))
abline(v = trueA, col="red" )

hist(chain[-(1:burnIn),2],nclass=30, main="Posterior of b", xlab="True value = red line")
abline(v = mean(chain[-(1:burnIn),2]))
abline(v = trueB, col="red" )

hist(chain[-(1:burnIn),3],nclass=30, main="Posterior of sd", xlab="True value = red line")
abline(v = mean(chain[-(1:burnIn),3]) )
abline(v = trueSd, col="red" )

plot(chain[-(1:burnIn),1], type = "l", xlab="True value = red line" , main = "Chain values of a" )
abline(h = trueA, col="red" )

plot(chain[-(1:burnIn),2], type = "l", xlab="True value = red line" , main = "Chain values of b")
abline(h = trueB, col="red" )

plot(chain[-(1:burnIn),3], type = "l", xlab="True value = red line" , main = "Chain values of sd")
abline(h = trueSd, col="red" )
 
# for comparison:
summary(lm(y~x))
```


reference:

https://theoreticalecology.wordpress.com/2010/09/17/metropolis-hastings-mcmc-in-r/
